---
title: 'Case Study 1: Birth Weight'
author: "Moriah Taylor, Faraz Yashar, AJ Eckmann, Audreya Metz, Naya Chen"
date: "9/4/2019"
output: html_document
---

```{r, message=FALSE, warning=FALSE, include=FALSE}
# Source for function: https://stackoverflow.com/questions/9341635/check-for-installed-packages-before-running-install-packages
pkgTest <- function(x)
  {
    if (!require(x,character.only = TRUE))
    {
      install.packages(x,repos = "http://cran.r-project.org", dep=TRUE)
        if(!require(x,character.only = TRUE)) stop("Package not found")
    }
  }

# Load any necessary packages here
pkgTest("tidyr")
pkgTest("dplyr")
pkgTest("broom")
pkgTest("hashmap")
pkgTest("MASS")
pkgTest("reshape2")
pkgTest("gridExtra")
pkgTest("ggplot2")
pkgTest("tidyverse")
pkgTest("quantreg")
library(quantreg)
library(ggplot2)
library(broom)
library(hashmap)
library(MASS)
library(reshape2)
library(gridExtra)
library(tidyverse)
library(tidyr)
library(dplyr)
set.seed(20190908)


#
# Helper functions
#

# Are we knitting?
ARE_KNITTING = isTRUE(getOption('knitr.in.progress'))

mean_sd_txt = function(x, digits = 0) {
  paste0(round(mean(x), digits = digits), "±", round(sd(x), digits = digits))
}

# function to convert numeric values to two-level categorical variables
# based on threshold
transform = function(data, cols, threshold) {
  for (i in 1:length(cols)) {
    col = cols[i]
    data[paste0(col, 'BIN')] = as.factor(
      case_when(
        data[col] <= threshold ~ 0,
        data[col] > threshold ~ 1
      )
    )
  }
  return(data)
}


inspect_covariate_vs_bwgt = function(data, covariate) {
  # covariate variable needs to be quoted: e.q. quo(blah) or quote(blah)
  data %>% group_by(!!covariate) %>% summarise(weight = mean(BWTG), se = sd(BWTG), N = length(BWTG)) %>%
  ggplot(aes_string(x=quo_name(covariate), y='weight')) +
    geom_errorbar(aes(ymin=weight-se, ymax=weight+se, color = N), width=0.1) +
    geom_line() +
    geom_point(aes(color = N))
}


# Plots a correlation for the given data
# requires ggplot, reshape2
plot_correlation_matrix = function(data) {
  cormat = cor(data)
  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
  upper_tri <- get_upper_tri(cormat)
  upper_tri
  melted_cormat <- melt(upper_tri, na.rm = TRUE)

  ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
  theme(axis.title.x=element_blank()) +
  theme(axis.title.y=element_blank()) +
  coord_fixed()
}

rmse = function(actual, predicted) {
  error = actual - predicted
  sqrt(mean(error^2))
}
```

```{r, include = FALSE, cache=TRUE}
# read in birth data with NAs
o_data = read.csv("data/Yr1116Birth.csv", na.strings=c("9","99", "9999"))

# remove missing rows
birth_data = na.omit(o_data)

# convert to categorical variables
# SEX, CORES(residence), MRACER(race), MHISP(Hispanic origin)
birth_data = birth_data %>% mutate(
  YOB = as.factor(YOB),
  SEX = as.factor(SEX),
  CORES = as.factor(CORES),
  MRACER = as.factor(MRACER),
  MHISP = as.factor(MHISP)
)

# convert average #cig to binary: smoke or not smoke
birth_data = transform(birth_data, c("CIGPN", "CIGFN", "CIGSN", "CIGLN"), 0)
# convert #parity to binary: pregnant before or never
birth_data = transform(birth_data, c("PARITY"), 1)

birth_data$SMOKER = (as.numeric(birth_data$CIGPNBIN) - 1) | # Before pregnancy
  (as.numeric(birth_data$CIGPNBIN) - 1) | # First trimester
  (as.numeric(birth_data$CIGPNBIN) - 1) | # Second trimester
  (as.numeric(birth_data$CIGPNBIN) - 1)   # Last trimester

# Combine all low frequency groups to other
birth_data$RACE = as.character(birth_data$MRACER)
birth_data$RACE[birth_data$RACE == "0"] = "O"
birth_data$RACE[birth_data$RACE == "1"] = "W"
birth_data$RACE[birth_data$RACE == "2"] = "B"
birth_data$RACE[birth_data$RACE == 3] = "O" # Native americans -> other
birth_data$RACE[birth_data$RACE == 4] = "A"
birth_data$RACE[birth_data$RACE == 5] = "A"
birth_data$RACE[birth_data$RACE == 6] = "A"
birth_data$RACE[birth_data$RACE == 7] = "A"
birth_data$RACE[birth_data$RACE == 8] = "A"
birth_data$RACE = as.factor(birth_data$RACE)
birth_data$RACE = relevel(birth_data$RACE, ref="W")

# 0 - Other non-White
# 1 - White
# 2 - Black or African American
# 3 - American Indian or Alaska Native
# 4 - Chinese
# 5 - Japanese
# 6 - Native Hawaiian
# 7 - Filipino
# 8 - Other Asian

# Separate hispanic and nonhispanic
birth_data$HISP = as.character(birth_data$MHISP)
birth_data$HISP[birth_data$HISP == "C"] = "Y"
birth_data$HISP[birth_data$HISP == "M"] = "Y"
birth_data$HISP[birth_data$HISP == "N"] = "N"
birth_data$HISP[birth_data$HISP == "O"] = "Y"
birth_data$HISP[birth_data$HISP == "P"] = "Y"
birth_data$HISP[birth_data$HISP == "S"] = "Y"
birth_data$HISP[birth_data$HISP == "U"] = "N"
birth_data$HISP = birth_data$HISP == "Y"
```

## Exploratory Analysis and Transformations

A normal pregnancy ranges from 38 to 42 weeks. (Source: https://medlineplus.gov/ency/article/002367.htm)

We note a gestational age nearly double that time at 80. We will remove the entry with that value since we consider it a clerical error.

```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
# remove the data point with more than 80 weeks gestational age 
birth_data = birth_data %>% filter(GEST < 50)
if (ARE_KNITTING) {
  birth_data_subset = birth_data
} else {
  birth_data_subset = birth_data[sample(nrow(birth_data), 100000),] # floor(nrow(birth_data) / 20)  
}
```


```{r echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
GEST_means = birth_data_subset %>% group_by(GEST) %>% summarise(weights = mean(BWTG), se = sd(BWTG)) %>% rename(BWTG = weights)

p1 <- ggplot(birth_data_subset,  aes(GEST, BWTG)) +
      geom_point(alpha =0.05) +
      geom_line(aes(GEST, BWTG), data = GEST_means, color = "red", alpha = 0.5) + 
      geom_errorbar(data = GEST_means, aes(GEST, BWTG, ymin=BWTG - se, ymax=BWTG + se), width=0.1, color = "red", alpha = 0.5) +
      ggtitle("Birth Weight vs. Gestational Age")

MAGE_means = birth_data_subset %>% group_by(MAGE) %>% summarise(weights = mean(BWTG), se = sd(BWTG)) %>% rename(BWTG = weights)

p2 <- ggplot(birth_data_subset, aes(MAGE, BWTG)) +
      geom_point(alpha = 0.05) +
      geom_line(aes(MAGE, BWTG), data = MAGE_means, color = "red", alpha = 0.5) + 
      geom_errorbar(data = MAGE_means, aes(MAGE, BWTG, ymin=BWTG - se, ymax=BWTG + se), width=0.1, color = "red", alpha = 0.5) +
      ggtitle("Birth Weight vs. Maternal Age")

PARITY_means = birth_data_subset %>% group_by(PARITY) %>% summarise(weights = mean(BWTG), se = sd(BWTG)) %>% rename(BWTG = weights)

p3 <- ggplot(birth_data_subset, aes(PARITY, BWTG)) +
      geom_point(alpha = 0.05) +
      geom_line(aes(PARITY, BWTG), data = PARITY_means, color = "red", alpha = 0.5) + 
      geom_errorbar(data = PARITY_means, aes(PARITY, BWTG, ymin=BWTG - se, ymax=BWTG + se), width=0.1, color = "red", alpha = 0.5) +
      ggtitle("Mean Weight by Parity")

p4 <- inspect_covariate_vs_bwgt(birth_data_subset, quote(PLUR)) +
      ggtitle("Mean Weight per Plurality") +
      theme(legend.position  = "none")

p5 <- ggplot(birth_data_subset, aes(MRACER, BWTG)) +
      geom_violin(aes(color = MRACER)) +
      stat_summary(fun.data=mean_sdl, mult=1, geom="pointrange") +
      geom_hline(yintercept = mean(birth_data_subset$BWTG), color = "red", alpha = 0.5) +
      ggtitle("Birth Weight vs. Race") + theme(legend.position = "none")

p6 <- ggplot(birth_data_subset, aes(MHISP, BWTG)) +
      geom_violin(aes(color = MHISP)) +
      stat_summary(fun.data=mean_sdl, mult=1, geom="pointrange") +
      geom_hline(yintercept = mean(birth_data_subset$BWTG), color = "red", alpha = 0.5) + 
      ggtitle("Birth Weight vs. Hispanic") + theme(legend.position = "none")

p7 <- ggplot(birth_data_subset, aes(YOB, BWTG)) +
      geom_violin(aes(color = YOB)) +
      stat_summary(fun.data=mean_sdl, mult=1, geom="pointrange") +
      ggtitle("Birth Weight by Year") +
      geom_hline(yintercept = mean(birth_data_subset$BWTG), color = "red", alpha = 0.5) + 
      theme(legend.position = "none")

p8 <- inspect_covariate_vs_bwgt(birth_data_subset, quote(CORES)) +
      theme(axis.text.x=element_blank()) +
      geom_hline(yintercept = mean(birth_data_subset$BWTG), color = "red", alpha = 0.5) + 
      ggtitle("Mean Weight by County") + theme(legend.position = "none")

p9 <- inspect_covariate_vs_bwgt(birth_data_subset, quote(CIGPN)) +
      geom_point(alpha = 0.1) +
      ggtitle("Mean Weight vs. Smoke Pre")

p10 <- inspect_covariate_vs_bwgt(birth_data_subset, quote(CIGFN)) +
      geom_point(alpha = 0.1) +
      ggtitle("Mean Weight vs. Smoke 1st Tri") + theme(legend.position = "none")

p11 <- inspect_covariate_vs_bwgt(birth_data_subset, quote(CIGSN)) +
      geom_point(alpha = 0.1) +
      ggtitle("Mean Weight by Smoke 2nd Tri") + theme(legend.position = "none")

p12 <- inspect_covariate_vs_bwgt(birth_data_subset, quote(CIGLN)) +
      geom_point(alpha = 0.1) +
      ggtitle("Mean Weight by Smoke 3nd Tri") + theme(legend.position = "none")
```

First we begin by inspecting the relationships between birth weight gestational age, maternal age, pairty, and plurality.

```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
grid.arrange(p1, p2, p3, p4, ncol=2)
```

We note that gestational age has a decent amount of curvature which might be helped with a quadratic term. This however would affect interpretability.

Maternal age seems to have little effect outside of the tails (e.g. greater than 45 years of age). Introduction of a nonlinearity would appear helpful here. Parity on birth weight seems to have a small effect. Plurality has a very pronounced effect.

We now inspect the relationship of weight with race, hispanic status, year of birth and county.

```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
grid.arrange(p5, p6, p7, p8, ncol=2)
```

Note the horizontal bar plotted is the average birthweight.

Inspecting race, certain race covariates appear to deviate signficantly from the mean, suggesting that this is a covariate of interest. Reading trough the literature, it appears that in another study of gestational variation published in the International Journal of Epidemiology entitled "Does gestation vary by ethnic group? A London-based study of over 122,000 pregnancies with spontaneous onset of labour" (https://academic.oup.com/ije/article/33/1/107/668109) categorization of european, african, and asian ancestry suffices.

```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
race_freqs = data.frame(table(birth_data$MRACER)) %>% gather(MRACER, value, -Var1) %>% 
    spread(Var1, value)
knitr::kable(race_freqs)
```

With this as prior knowledge, we'll consider grouping together all of the racial categories of asian ancestry to simplify interpreation.

Due to their small size, the granular subsets of hispanic ethnicity may not be very informative. Investigating the covariate directly, we note the majority reported as not hispanic.

```{r echo=FALSE, cache=TRUE}
hisp_freqs = data.frame(table(birth_data$MHISP)) %>% gather(MHISP, value, -Var1) %>% 
    spread(Var1, value)
knitr::kable(hisp_freqs)
```

As such, we will consider a transformation to turn hispanic ethnicity into a binary covariate.

Certain counties appear to have significant deviation from the mean. Since county is a proxy for socioeconomic status, it would be worthwhile to link counties through census data to useful socioeconomic indicator such as median income. This would reduce the number of covariates significantly. This would require the addition of another dataset to our analysis, so for the purpose of this study we will choose to ignore the CORES variable.

Inspecting year of bith more closely, we see little variation in any of the covariates year over year.

```{r, echo=FALSE, cache=TRUE}
birth_data %>% group_by(YOB) %>% summarize(
  GEST = mean_sd_txt(GEST, 2),
  BWTG = mean_sd_txt(BWTG, 2),
  SMOKER = round(mean(SMOKER == TRUE), 2),
  SEX = round(mean(SEX == 1), 2),
  MAGE = mean_sd_txt(MAGE, 1),
  PLUR = mean_sd_txt(PLUR, 1),
  PARITY = mean_sd_txt(PARITY, 1),
  ASIAN = round(mean(RACE == 'A'), 2),
  BLACK = round(mean(RACE == 'B'), 2),
  WHITE = round(mean(RACE == 'W'), 2),
  OTHER = round(mean(RACE == 'O'), 2),
  HISP = round(mean(HISP == 1), 3)
) %>% knitr::kable()
```

As such, we believe that there is not a significant time-related trend and that year of birth is potentially not a covariate of interest.

We now inspect the smoking variables.

```{r echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
grid.arrange(p9, p10, p11, p12, ncol=2)
```

Note in most cases we observe an initial drop from zero followed by wild fluctuations as the number of cigarettes smoked per day increases. The vast majority of mothers are non smokers throughout each phase of pregnancy.

 - `r paste0(round(mean(birth_data$CIGPN == 0), 3) * 100, '%')` of women don't smoke before pregancy.
 - `r paste0(round(mean(birth_data$CIGFN == 0), 3) * 100, '%')` of women don't smoke during the first trimester.
 - `r paste0(round(mean(birth_data$CIGSN == 0), 3) * 100, '%')` of women don't smoke during the second trimester.
 - `r paste0(round(mean(birth_data$CIGLN == 0), 3) * 100, '%')` of women don't smoke during the last trimester.
 
In addition, a correlation matrix revealed heavy correlations between all of the smoking variables.

```{r echo=FALSE}
plot_correlation_matrix(na.omit(o_data) %>% dplyr::select(-MHISP))
```

This lead us to beleive that binarizing each of the smoking covariates or combining them into a single covariate would make sense.

## Model Selection

Based on the exploratory data analysis we devised the following suite of models to train and evaluate:

```{r, echo=FALSE, cache=FALSE}
models = list(
  # All covariates included without transformation
  full = "BWTG ~ YOB + SEX + CORES + CIGPN  + CIGSN + CIGLN + GEST + PLUR + MAGE + MRACER + MHISP + PARITY",
  
  # Full model with YOB and CORES dropped to simplify interpretation
  dyob_dcores = "BWTG ~ SEX + CIGPN + CIGFN + CIGSN + CIGLN + GEST + PLUR + MAGE + MRACER + MHISP + PARITY",
  
  #
  # All the models below have YOB and CORES dropped
  #
  
  # With smoking (per term) binarized
  bsmokes = "BWTG ~ SEX + CIGPNBIN + CIGFNBIN + CIGSNBIN + CIGLNBIN + GEST + PLUR + MAGE + MRACER + MHISP + PARITY",
  
  # Binarize smoking (per term), parity
  bsmokes_bpar = "BWTG ~ SEX + CIGPNBIN + CIGFNBIN + CIGSNBIN + CIGLNBIN + GEST + PLUR + MAGE + MRACER + MHISP + PARITYBIN",
  
  # Binarize smoking (per term), parity, hisp
  bsmokes_bpar_bhisp = "BWTG ~ SEX + CIGPNBIN + CIGFNBIN + CIGSNBIN + CIGLNBIN + GEST + PLUR + MAGE + MRACER + HISP + PARITYBIN",
  
  # Binarize smoking (per term), binarize party, recategorize race, binarize hispanic
  bsmokes_bpar_crace_bhisp = "BWTG ~ SEX + CIGPNBIN + CIGFNBIN + CIGSNBIN + CIGLNBIN + GEST + PLUR + MAGE + RACE + HISP + PARITYBIN",
  
  # Binarize smoking (per term), binarize party, drop hispanic
  bsmokes_bpar_crace_dhisp = "BWTG ~ SEX + CIGPNBIN + CIGFNBIN + CIGSNBIN + CIGLNBIN + GEST + PLUR + MAGE + RACE + PARITYBIN",
  
  # Binarize smoking categories, square gestation
  bsmokes_bpar_crace_dhisp_gest2 = "BWTG ~ SEX + CIGPNBIN + CIGFNBIN + CIGSNBIN + CIGLNBIN + GEST^2 + PLUR + MAGE + RACE + PARITYBIN",
  
  # Binarize smoking categories, square gestation, square mage
  bsmokes_bpar_crace_dhisp_mage2 = "BWTG ~ SEX + CIGPNBIN + CIGFNBIN + CIGSNBIN + CIGLNBIN + GEST + PLUR + MAGE^2 + RACE + PARITYBIN",
  
  # Binarize smoking into a single variable
  bsmoke = "BWTG ~ SEX + SMOKER + GEST + PLUR + MAGE + MRACER + MHISP + PARITY",

  # Simplest model for interpretation
  simple = "BWTG ~ SEX + SMOKER + GEST + PLUR + MAGE + RACE + HISP + PARITYBIN",
  
  #
  # Below are log transformations of some of the above models
  #
  
  full_log = "log(BWTG) ~ YOB + SEX + CORES + CIGPN  + CIGSN + CIGLN + GEST + PLUR + MAGE + MRACER + MHISP + PARITY",
  dyob_dcores_log = "log(BWTG) ~ SEX + CIGPN + CIGFN + CIGSN + CIGLN + GEST + PLUR + MAGE + MRACER + MHISP + PARITY",
  bsmokes_log = "log(BWTG) ~ SEX + CIGPNBIN + CIGFNBIN + CIGSNBIN + CIGLNBIN + GEST + PLUR + MAGE + MRACER + MHISP + PARITY",
  bsmoke_log = "log(BWTG) ~ SEX + SMOKER + GEST + PLUR + MAGE + MRACER + MHISP + PARITY",
  simple_log = "log(BWTG) ~ SEX + SMOKER + GEST + PLUR + MAGE + RACE + HISP + PARITY"
)
```

```{r, include=FALSE, cache=TRUE}
N = nrow(birth_data)
folds = 5
holdouts = split(sample(1:N), 1:folds)

results = data.frame(melt(models)[,c(2,1)]) %>% rename(name = L1, formula = value) %>% mutate(formula = as.character(formula))
results$rsquared = 0
results$rmse = 0
results$finished = FALSE
tibble(results)

for (model_name in names(models)) {
  if (results[results$name == model_name,]$finished) {
    next()
  }
  formula = models[[model_name]]
  print(paste0("Running model ", formula))
  
  rsquareds = c()
  rmses = c()
  
  for (i in 1:folds) {
    print(paste0("-- Fold ", i))
    test_indices = holdouts[[i]]
    
    test = birth_data[test_indices,]
    train = birth_data[-test_indices,]
  
    model = lm(formula, data = train)

    rsquared = summary(model)$r.squared
    
    preds = predict(model, test)
    
    if (grepl("log", formula)) {
      err = rmse(test$BWTG, exp(preds))
    } else {
      err = rmse(test$BWTG, preds)  
    }
    
    rmses = c(rmses, err)
    rsquareds = c(rsquareds, rsquared)
  }
  
  results[results$name == model_name,]$rsquared = mean(rsquareds)
  results[results$name == model_name,]$rmse = mean(rmses)
  results[results$name == model_name,]$finished = TRUE
  
  print(paste0("Finished model ", formula))
}
```

With 5-fold cross validation, we see the following results for all of the models tested:

```{r, echo= FALSE, cache= TRUE}
formatted_results = results %>% dplyr::select(-finished) %>% mutate(rsquared = round(rsquared, 4), rmse = round(rmse, 2)) %>% arrange(rmse, -rsquared)
formatted_results %>% knitr::kable()
```

Investigating the models above, we see that log transforming the outcome improves fit greatly, however this is clearly a result of overfitting as the RMS errors are the highest for those transformations. Binarization increased performance broadly. Introducing quadratic terms did nothing. Of the models with the lowest RMSE (within 0.1 from the smallest), we see that the model which binarizes the smoking variables, combines the races as mentioned prior, and drops the hispanic term (bsmokes_bpar_crace_dhisp) performs extremely well when compared against the other models of similar performance. In our view, this model has the optimal blend of simplicity, interpretability and performance.

Now that we have selected our covariates, we'll build our final OLS model and compare it against a robust regression.

## Model Evaluation

We now train the model across the entire data set and evaluate it's performance.

```{r, echo=FALSE, cache=TRUE}
ols = lm(BWTG ~ SEX + CIGPNBIN + CIGFNBIN + CIGSNBIN + CIGLNBIN + GEST + PLUR + MAGE + RACE + PARITYBIN, data = birth_data[1:10000,])
summary(ols)
par(mfrow=c(2,2))
plot(ols)
```

In examining the plots, it's clear that the assumption of heteroskedasticity is violated.
In addition, the normality assumption is somewhat violated since the residuals appear to have fatter tails.
We'll now move on to robust regression to see if there is a significant improvement.

```{r, echo=FALSE, cache=TRUE}
robust = rlm(BWTG ~ SEX + CIGPNBIN + CIGFNBIN + CIGSNBIN + CIGLNBIN + GEST + PLUR + MAGE + RACE + PARITYBIN, data = birth_data[1:10000,])
summary(robust)
par(mfrow=c(2,2))
plot(robust)
```

In analyzing the robust regression, we see little improvement. Furthermore, we see little change in the effect sizes of the covariates. As such, we choose our OLS model for reporting.

## Quantile Regression

### Full Model Based on OLS
```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
quant1 = rq(BWTG ~ SEX + CIGPNBIN + CIGFNBIN + CIGSNBIN + CIGLNBIN + GEST + PLUR + MAGE + RACE + PARITYBIN, data = birth_data[1:10000,], tau=c(.05, .25, .5, .75, .95))
summary(quant1)
plot(quant1)
```

### Model Without Smoking Status
```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
quant2 = rq(BWTG ~ SEX + GEST + PLUR + MAGE + RACE + PARITYBIN, data = birth_data[1:10000,], tau=c(.05, .25, .5, .75, .95))
summary(quant2)
plot(quant2)
```


### Model With Smoking Binary
```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
quant3 = rq(BWTG ~ SEX + SMOKER + GEST + PLUR + MAGE + RACE + PARITYBIN, data = birth_data[1:10000,], tau=c(.05, .25, .5, .75, .95))
summary(quant3)
plot(quant3)
```

## Interpretation and Conclusion

TODO: Interpretation of OLS model for RACE = W, SEX = 1, MAGE = 28, GEST = 40, PLUR = 1, PARITY = 0

TODO: Interpretation of quantile model
